Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-02-04 paper/manuscript.tex
Gary walk through

Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: 
\begin{itemize}
\item a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{D})\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; 
\item the entire ligand can be alchemical transferred from protein to solvent in an absolute binding free energy calculation (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{E})~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; 
\item and small alchemical modifications can be made between chemically related ligands to estimate relative differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{F})~\cite{wang2015accurate,mey2016blinded,gapsys2020large}.
\end{itemize}
%

@@

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
What can be expected from this Best Practice guide is a good starting point for new practitioners and a reference for experience practitioners. For this propose we provide a convenient checklist (Section~\ref{sec:checklist}) to help ensure all calculations meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.
What we assume from a novice practitioner is a moderate experience with molecular simulation packages and concepts. Furthermore basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association are essential. If you feel unfamiliar with some of these concepts, good starting points can be found in these references~\cite{braun2019best, grossfield2018best, klimovich2015guidelines, shirts2012best}. While reading this Best Practice document, it is important to bare in mind \emph{this is not a review} of all free energy calculation methods at the cutting edge of current research.
Instead the guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
\end{itemize}
%

@@

\paragraph{Simulating binding events is computationally expensive}
While simulating binding events has been used to estimate binding affinities~\cite{jong_determining_2011,pan2017quantitative} or to get insights into the binding pathways and kinetics of receptor-ligand systems~\cite{teo_adaptive_2016,votapka_seekr_2017,doerr_--fly_2014,plattner_protein_2015,dixon_predicting_2018}, the computational cost of these calculations is usually dominated by the rate of dissociation, which can be on the $\mu$s scale even for millimolar binders~\cite{pan2017quantitative} and reaches the ms or s scale for a typical drug~\cite{basavapathruni_conformational_2012,hyre_cooperative_2006}.
Depending on system size and simulation settings, common molecular dynamics software packages can reach a few hundreds of ns/day using currently available high-end GPUs~\cite{eastman_openmm_2017,kutzner_more_2019}, making these type of calculations uninteresting at pharmaceutical drug discovery timescales.
Other methods compute the free energy of binding by building potential of mean force profiles along a reaction coordinate~\cite{woo2005calculationa,velezvega_overcoming_2013,limongelli_funnel_2013,heinzelmann2017attachpullrelease}, but these methods require prior knowledge of a high-probability binding pathway, which is not easily available, especially in the prospective scenarios typical of the drug development process.

@@

In many cases, the quantity of interest is the change in binding affinity when a compound $A$ is mutated to compound $B$ (e.g., by modifying one the drug scaffold's substituents), which, by using Eq.~\ref{eq:binding-free-energy-from-bound-unbound-probability-ratio} and \ref{eq:bound-unbound-probability-ratio} is given by
\begin{equation}\label{eq:delta-delta-G-physical}
\begin{split}
    \Delta \Delta G_{\mathrm{bind}, AB} &= \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A} \\
    &\approx -k_BT \left( \ln \frac{Z(RB)}{Z(R+B)} - \ln \frac{Z(RA)}{Z(R+A)} \right) \, .
\end{split}
\end{equation}
Note that the terms involving the standard concentration cancel out when we assume that the volume is identical for $A$ and $B$.
Predictions of $\Delta \Delta G_{\mathrm{bind}, AB}$ with non-alchemical methods generally require long simulations of both ligands, possibly through different binding pathways.
Alchemical relative free energy calculations avoid the need to simulate binding and unbinding events by making use of the fact that the free energy is a state function and exploiting the thermodynamic cycle illustrated in Figure~\ref{fig:fig_binding_thermodynamic_cycle}.
This is apparent after rewriting Eq.~\ref{eq:delta-delta-G-physical} as
\begin{equation}\label{eq:delta-delta-G-alchemical}
\begin{split}
    \Delta \Delta G_{\mathrm{bind}, AB} &\approx -k_BT \left( \ln \frac{Z(RB)}{Z(RA)} - \ln \frac{Z(R+B)}{Z(R+A)} \right) \\
    &= -k_BT \left( \ln \frac{Z(RB)}{Z(RA)} - \ln \frac{Z(B)}{Z(A)} \right) \\
    &= \Delta G_{\mathrm{bound}} - \Delta G_{\mathrm{unbound}} \, ,
\end{split}
\end{equation}
where $\Delta G_{\mathrm{bound/unbound}}$ is the free energy of mutating $A$ to $B$ in the bound/unbound state.
Eq.~\ref{eq:delta-delta-G-alchemical} and Figure~\ref{fig:fig_binding_thermodynamic_cycle} tell us that the difference in free energy of binding between toluene ($A$) and benzyl alcohol ($B$) can be computed by running two independent calculations estimating the free energy cost of mutating $A$ into $B$ in the binding pocket ($\Delta G_{\mathrm{bound}}$) and in solvent ($\Delta G_{\mathrm{unbound}}$), saving us the need to simulate the physical binding process of the two compounds.
In particular, the second line of Eq.~\ref{eq:delta-delta-G-alchemical} is a consequence of $\Delta G_{\mathrm{unbound}}$ being independent of the presence of the receptor in the simulation box as the definition of the unbound state assumes receptor and ligand to be at a sufficient distance for them to have no energetic interactions.
Note that, when $A$ and $B$ have different number of atoms, Eq.~\ref{eq:delta-delta-G-alchemical} implies the presence of a factor having units of volume entering both logarithms, which require unitless arguments.
The factor's value is, however, inconsequential as it cancels out in practice.

@@

In practice, the mutation of $A$ to $B$ in carried out by introducing one or more parameters $\vec{\lambda}$ controlling the potential energy function $U(\vec{q};\vec{\lambda})$ such that the potential of compounds $A$ and $B$ is recovered at two particular values $\vec{\lambda}_A$ and $\vec{\lambda}_B$.
Briefly, this is achieved by simulating a ``chimera'' molecule composed by enough atoms to represent both $A$ and $B$.
A subset of the energetic terms in $U(\vec{q};\vec{\lambda})$ is then modulated by $\vec{\lambda}$ so that at $\vec{\lambda}_A$, the atoms that form molecule $A$ are activated and those belonging exclusively to $B$ are non-interacting ``dummy atoms'', while the opposite occur at $\lambda_B$ (see Section~\ref{sec:relative-fe-protocol} for details).

We can rigorously account for fluctuations in other thermodynamic parameters such as changes in volume $V$ when simulating at constant pressure $p$ or changes in number of molecules $N_i$ of species $i$ at constant chemical potential $\mu_i$ (e.g., number of waters or ions) by introducing the \textit{reduced potential}~\cite{shirts2008statisticallya}
\begin{equation}\label{eq:reduced-potential}
u(\vec{q};\vec{\lambda}) \equiv \beta \left[ U(\vec{q};\vec{\lambda}) + p \, V(\vec{q}) + \sum_i \mu_i \, N_i(\vec{q}) + \cdots \right] \, .
\end{equation}
Here, the collection of thermodynamic and alchemical parameters $\{\beta, \vec{\lambda}, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
In the context of alchemical calculations, in which the thermodynamic states vary only in their value of $\vec{\lambda}$, these are also referred to as \emph{alchemical states}.
The free energy of mutating $A$ to $B$ in any environment (e.g., binding site, solvent) can then be computed as
\begin{equation}\label{eq:delta-G-env}
    \Delta G_{\mathrm{env}} = - k_BT \ln \frac{Z(\vec{\lambda}_B)}{Z(\vec{\lambda}_A)} = - k_BT \ln \frac{\int_{\Gamma_{\mathrm{env}}} \exp\left( u(\vec{q}; \vec{\lambda}_B) \right) \, d\vec{q}}{\int_{\Gamma_{\mathrm{env}}} \exp\left( u(\vec{q}; \vec{\lambda}_A) \right) \, d\vec{q}} \, .
\end{equation}
While it is generally not feasible to compute the two partition functions $Z(\vec{\lambda})$, several estimators have been devised to robustly estimate the ratio of partition functions in Eq.~\ref{eq:delta-G-env} (see Section~\ref{subsec:estimators}) from a set of conformations usually collected with MD simulations from the thermodynamic states defined at $\vec{\lambda}_A$ and $\vec{\lambda}_B$.

\paragraph{Why do alchemical calculations need unphysical intermediate states?}
While it is theoretically possible to estimate the ratio of partition functions from samples collected only at states $\vec{\lambda}_A$ and $\vec{\lambda}_B$, the efficiency of the free energy estimators rapidly decreases with the phase-space overlap between the two states~\cite{wu2005phasespaceb,wu2005phasespacec}.
Roughly, the phase-space overlap between two thermodynamic states measures the degree to which high-probability conformations (i.e., those with very negative potential energy) in one state are high-probability conformations in the other state as well (see Section~\ref{sec:are-they-good}).

To solve the of having poor overlap between the states of interest, multiple intermediate alchemical states are introduced at values $\vec{\lambda}_A = \vec{\lambda}_0, \vec{\lambda}_1, \cdots, \vec{\lambda}_K = \vec{\lambda}_B$ so that each pair of consecutive states $\vec{\lambda}_k, \vec{\lambda}_{k+1}$ share good overlap.
Each intermediate state models a ligand that is neither $A$ nor $B$ but a mix of the two.
Many estimators (e.g., exponential reweighting (EXP)~\cite{zwanzig1954hightemperature} and Bennet acceptance ratio (BAR)~\cite{bennett1976efficient,shirts2003equilibriuma}) can then be used to compute the free energy as
\begin{equation}
    \Delta G_{\mathrm{env}} = k_BT \sum_{k=0}^{K-1} \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1})
\end{equation}
from samples collected at all the alchemical states $\{\vec{\lambda}_k \}$, where $\Delta f$ is the \emph{unitless free energy difference}
\begin{equation}
    \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1}) = f(\vec{\lambda}_{k+1}) - f(\vec{\lambda}_k) = - \ln \frac{Z(\vec{\lambda}_{k+1})}{Z(\vec{\lambda}_k)} \, .
\end{equation}
While this strategy usually results in sampling thermodynamic states whose Boltzmann distributions is very similar, thus collecting information that is to some degree redundant, some estimators such as the Multistate Bennett acceptance ratio (MBAR)~\cite{shirts2008statisticallya} can exploit similarities between states to improve the precision of the estimates by using the conformations sampled in all alchemical states $\{\vec{\lambda}_k \}$ to compute the free energy difference $\Delta f(\vec{\lambda}_i, \vec{\lambda}_{j})$ between any pair of states $i,j$  (see Section~\ref{subsec:estimators}).

@@

To ensure modeling matches experiment, we also need to ensure that we simulate the same system -- which requires understanding what protein construct is used in the bioassay. For instance, does the X-ray structure that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer etc)~\cite{perez-benito2018predicting}? Also, were certain co-factors or partner proteins required in the bioassay? 
%
\subsection{Is my binding mode accurate?}
As also mentioned, good performance of alchemical FE calculations requires an accurate representation of the ligand binding mode, which often is estimated based on X-ray crystallography. If more than one structure is available, the modeler should pay attention to choose the most suitable. The quality of the structure can be a concern, and the reader is referred to work of Warren et al. for a detailed discussion of choosing optimal structures for structure-based modeling~\cite{warren2012essential}.

It is also useful to study the structure activity relationship and understand the expected impact in the binding site, whether side chain movement in the protein will be required, and whether there evidence of this in any alternative X-ray structures of the same protein. Often, only one protein and water configuration is used for a series of alchemical FE calculations, so this needs to be capable of accommodating the smallest through to largest ligands in a way that allows stable and well behaved simulations. This can provide a practical limit on the alchemical changes that are feasible, though a simple work-around can be to separate compounds into sub-series for different calculations.

@@

Finally, the role of water in ligand binding is not always well understood and it can be crucial to capture the changes in binding site solvation during ligand binding. Can crystallographic waters be retained? Do they clash with some of the larger ligands used in the FEP? If so, how will the water be incorporated for smaller ligands that require the water – consider pre-solvation methods such as GCMC steps~\cite{michel2010prediction}. Other tools such as WaterMap or open source equivalents (SSTmap, others) can be used to define water structure for systems with no experimental evidence of water sites. Before launching large numbers of alchemical free energy calculations it is recommended to test the system using classical MD simulations and limited numbers of alchemical perturbations. Metrics such as ligand and protein RMSD and RMSF can be inspected, along with visual inspection of simulations, to ensure the system is stable and likely to be suitable for FE calculations.

@@

\paragraph{How to deal with different binding modes?}
Generally, binding modes within congeneric series are conserved~\cite{wacker2010conserved}, however, exceptions exist~\cite{brandt2011congeneric,nazare2005probing}, as discussed in more detail in Section~\ref{sec:multiple_binding_modes}. Certain  functional groups may be particularly prone to this due to symmetries or near symmetries. One such issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome bias due to the starting configuration. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.
%
\paragraph{Have you considered sterioisomers and enationmers?}
Congeneric series can contain stereoisomers or enantiomers which can bind very differently, resulting in large errors if treated incorrectly. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one stereoisomer/enantiomer is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces potential for larger errors compared to experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Nevertheless, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer. Additionally, unexpected changes in what stereoisomer binds experimentally, if they occur, could pose significant challenges for modelling efforts.
%
\paragraph{Conserved binding site waters can play an important role in binding free energies}
Binding site water molecules may form water mediated protein-ligand interactions which can pose challenges whenever exchange with bulk water is slow compared to simulation timescales. This happens typically in buried binding sites. Overlaying multiple protein X-ray structures can identify conserved or additional water molecules that can be useful to include in calculations. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples, see HSP90 which formed part of the D3R grand challenge 2015~\cite{mey2016blinded}, A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.
%
\paragraph{How to align congeneric series}
Input coordinates for a congeneric series may be generated by docking calculations, or by ligand alignment using maximum common substructure algorithms. The latter tends to produce alignments that are more conserved and more consistent free energy changes across a dataset, but will struggle to yield reasonable results for relative binding free energy calculations that involve a significant binding mode rearrangement. This may also lead to steric clashes with the receptor coordinates of the reference ligand if structural rearrangements are needed to accommodate different members of the congeneric series. Small steric clashes may be resolved during subsequent simulation equilibration prior to data collection, but there is a risk that the complex relaxes to an alternative metastable state.

@@

In this case the common substructure is the benzene ring, though perhaps may be larger depending on how it is defined, as we discuss below.

@@

Rigorously, this process comprises of a maximal common substructure (MCSS) search of the molecules involved to identify the common substructure -- though the parameters of the MCSS search will differ depending on whether single or dual topology calculations are planned.

@@

Some approaches have attempted to address this~\cite{clark2019relative} but a general solution is not yet in mainstream use.

@@

Tthe use of restraints can therefore dramatically improve sampling as interactions are weakened and removed.

@@

Alternatively, a flat-bottom restraint can be activate also in the decoupled state as long as the second restraint (e.g., a harmonic or Boresch restraint) prevent the ligand to hit the wall of the flat-bottom potential~\cite{rizzi2019sampl6}.

@@

Specifically, finite-size effects can introduce profound artifacts into computed binding free energies~\cite{}, in part because typical schemes for long-range electrostatics (including PME and reaction field) do not handle free energy contributions from such changes effectively or as they would be handled in a hypothetical macroscopic bulk solution~\cite{}.

@@

Many enhanced sampling techniques have been proposed to try and overcome barriers in free energy calculations and mitigate convergence to erroneous estimates~\cite{}. 

>>>>

Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: 
\begin{itemize}
\item a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{D})\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; 
\item the entire ligand can be alchemically transferred from protein to solvent in an absolute binding free energy calculation (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{E})~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; 
\item and small alchemical modifications can be made between chemically related ligands to estimate relative differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} \textbf{F})~\cite{wang2015accurate,mey2016blinded,gapsys2020large}.
\end{itemize}
%

@@

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
What can be expected from this Best Practice guide is a good starting point for new practitioners and a reference for experienced practitioners. For this propose we provide a convenient checklist (Section~\ref{sec:checklist}) to help ensure all calculations meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.
What we assume from a novice practitioner is a moderate experience with molecular simulation packages and concepts. Furthermore, basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association are essential. If you feel unfamiliar with some of these concepts, good starting points can be found in these references~\cite{braun2019best, grossfield2018best, klimovich2015guidelines, shirts2012best}. While reading this Best Practice document, it is important to bare in mind \emph{this is not a review} of all free energy calculation methods at the cutting edge of current research.
Instead the guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
\end{itemize}
%

@@

\paragraph{Simulating binding events is computationally expensive}
While simulating binding events has been used to estimate binding affinities~\cite{jong2011determining,pan2017quantitative} or to get insights into the binding pathways and kinetics of receptor-ligand systems~\cite{teo2016adaptive,votapka2017seekr,doerr2014onthefly,plattner2015protein,dixon2018predicting}, the computational cost of these calculations is usually dominated by the rate of dissociation, which can be on the {\textmu}s scale even for millimolar binders~\cite{pan2017quantitative} and reaches the ms or s scale for a typical drug~\cite{basavapathruni2012conformational,hyre2006cooperative}.
Depending on system size and simulation settings, common molecular dynamics software packages can reach a few hundreds of ns/day using currently available high-end GPUs~\cite{eastman2017openmm,kutzner2019more}, making these type of calculations uninteresting at pharmaceutical drug discovery timescales.
Other methods compute the free energy of binding by building potential of mean force profiles along a reaction coordinate~\cite{woo2005calculationa,velez-vega2013overcoming,limongelli2013funnel,heinzelmann2017attachpullrelease}, but these methods require prior knowledge of a high-probability binding pathway, which is not easily available, especially in the prospective scenarios typical of the drug development process.

@@

In many cases, the quantity of interest is the change in binding affinity when a compound $A$ is mutated to compound $B$ (e.g., by modifying one the drug scaffold's substituents, see Fig.~\ref{fig:fig_what_is_alchemy} \textbf{F}), which, by using Eq.~\ref{eq:binding-free-energy-from-bound-unbound-probability-ratio} and \ref{eq:bound-unbound-probability-ratio} is given by
\begin{equation}\label{eq:delta-delta-G-physical}
\begin{split}
    \Delta \Delta G_{\mathrm{bind}, AB} &= \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A} \\
    &\approx -k_BT \left( \ln \frac{Z(RB)}{Z(R+B)} - \ln \frac{Z(RA)}{Z(R+A)} \right) \, .
\end{split}
\end{equation}
Note that the terms involving the standard concentration cancel out when we assume that the volume is identical for $A$ and $B$.
Predictions of $\Delta \Delta G_{\mathrm{bind}, AB}$ with non-alchemical methods generally require long simulations of both ligands, possibly through different binding pathways.
Alchemical relative free energy calculations avoid the need to simulate binding and unbinding events by making use of the fact that the free energy is a state function and exploiting the thermodynamic cycle illustrated in Figure~\ref{fig:fig_binding_thermodynamic_cycle}.
This is apparent after rewriting Eq.~\ref{eq:delta-delta-G-physical} as
\begin{equation}\label{eq:delta-delta-G-alchemical}
\begin{split}
    \Delta \Delta G_{\mathrm{bind}, AB} &\approx -k_BT \left( \ln \frac{Z(RB)}{Z(RA)} - \ln \frac{Z(R+B)}{Z(R+A)} \right) \\
    &= -k_BT \left( \ln \frac{Z(RB)}{Z(RA)} - \ln \frac{Z(B)}{Z(A)} \right) \\
    &= \Delta G_{\mathrm{bound}} - \Delta G_{\mathrm{unbound}} \, ,
\end{split}
\end{equation}
where $\Delta G_{\mathrm{bound/unbound}}$ is the free energy of mutating $A$ to $B$ in the bound/unbound state.
Eq.~\ref{eq:delta-delta-G-alchemical} and Figure~\ref{fig:fig_binding_thermodynamic_cycle} tell us that the difference in free energy of binding between toluene ($A$) and benzyl alcohol ($B$) can be computed by running two independent calculations estimating the free energy cost of mutating $A$ into $B$ in the binding pocket ($\Delta G_{\mathrm{bound}}$) and in solvent ($\Delta G_{\mathrm{unbound}}$), saving us the need to simulate the physical binding process of the two compounds.
In particular, the second line of Eq.~\ref{eq:delta-delta-G-alchemical} is a consequence of $\Delta G_{\mathrm{unbound}}$ being independent of the presence of the receptor in the simulation box as the definition of the unbound state assumes receptor and ligand to be at a sufficient distance for them to have no energetic interactions.
Note that, when $A$ and $B$ have different number of atoms, Eq.~\ref{eq:delta-delta-G-alchemical} implies the presence of a factor having units of volume entering both logarithms, which require unitless arguments.
The factor's value is, however, inconsequential as it cancels out in practice.

@@

In practice, the mutation of $A$ to $B$ in carried out by introducing one or more parameters $\vec{\lambda}$ controlling the potential energy function $U(\vec{q};\vec{\lambda})$ such that the potential of compounds $A$ and $B$ is recovered at two particular values $\vec{\lambda}_A$ and $\vec{\lambda}_B$.
Briefly, this is achieved by simulating a ``chimera'' molecule composed of enough atoms to represent both $A$ and $B$.
A subset of the energetic terms in $U(\vec{q};\vec{\lambda})$ is then modulated by $\vec{\lambda}$ so that at $\vec{\lambda}_A$, the atoms that form molecule $A$ are activated and those belonging exclusively to $B$ are non-interacting ``dummy atoms'', while the opposite occur at $\lambda_B$ (see Section~\ref{sec:relative-fe-protocol} for details).

We can rigorously account for fluctuations in other thermodynamic parameters such as changes in volume $V$ when simulating at constant pressure $p$ or changes in number of molecules $N_i$ of species $i$ at constant chemical potential $\mu_i$ (e.g., number of waters or ions) by introducing the \textit{reduced potential}~\cite{shirts2008statisticallya}
\begin{equation}\label{eq:reduced-potential}
u(\vec{q};\vec{\lambda}) \equiv \beta \left[ U(\vec{q};\vec{\lambda}) + p \, V(\vec{q}) + \sum_i \mu_i \, N_i(\vec{q}) + \cdots \right] \, .
\end{equation}
Here, the collection of thermodynamic and alchemical parameters $\{\beta, \vec{\lambda}, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
In the context of alchemical calculations, in which the thermodynamic states vary only in their value of $\vec{\lambda}$, these are also referred to as \emph{alchemical states}.
The free energy of mutating $A$ to $B$ in any environment (e.g., binding site, solvent) can then be computed as
\begin{equation}\label{eq:delta-G-env}
    \Delta G_{\mathrm{env}} = - k_BT \ln \frac{Z(\vec{\lambda}_B)}{Z(\vec{\lambda}_A)} = - k_BT \ln \frac{\int_{\Gamma_{\mathrm{env}}} \exp\left( u(\vec{q}; \vec{\lambda}_B) \right) \, d\vec{q}}{\int_{\Gamma_{\mathrm{env}}} \exp\left( u(\vec{q}; \vec{\lambda}_A) \right) \, d\vec{q}} \, .
\end{equation}
While it is generally not feasible to compute the two partition functions $Z(\vec{\lambda})$, several estimators have been devised to robustly estimate the ratio of partition functions in Eq.~\ref{eq:delta-G-env} (see Section~\ref{subsec:estimators}) from a set of conformations usually collected with MD simulations from the thermodynamic states defined at $\vec{\lambda}_A$ and $\vec{\lambda}_B$ and intermediates thereof.

\paragraph{Why do alchemical calculations need unphysical intermediate states?}
While it is theoretically possible to estimate the ratio of partition functions from samples collected only at states $\vec{\lambda}_A$ and $\vec{\lambda}_B$, the efficiency of the free energy estimators rapidly decreases as the phase-space overlap between the two states also decreases~\cite{wu2005phasespaceb,wu2005phasespacec}.
Roughly, the phase-space overlap between two thermodynamic states measures the degree to which high-probability conformations (i.e., those with very negative potential energy) in one state are high-probability conformations in the other state as well (see Section~\ref{sec:are-they-good} and Fig.~\ref{fig:fig_what_is_lambda}).

To solve the problem of having poor overlap between the states of interest, multiple intermediate alchemical states are introduced at values $\vec{\lambda}_A = \vec{\lambda}_0, \vec{\lambda}_1, \cdots, \vec{\lambda}_K = \vec{\lambda}_B$ so that each pair of consecutive states $\vec{\lambda}_k, \vec{\lambda}_{k+1}$ share good overlap.
Each intermediate state models a ligand that is neither $A$ nor $B$ but a mix of the two.
Many estimators (e.g., exponential reweighting (EXP)~\cite{zwanzig1954hightemperature} and Bennet's acceptance ratio (BAR)~\cite{bennett1976efficient,shirts2003equilibriuma}) can then be used to compute the free energy as
\begin{equation}
    \Delta G_{\mathrm{env}} = k_BT \sum_{k=0}^{K-1} \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1})
\end{equation}
from samples collected at all the alchemical states $\{\vec{\lambda}_k \}$, where $\Delta f$ is the \emph{unitless free energy difference}
\begin{equation}
    \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1}) = f(\vec{\lambda}_{k+1}) - f(\vec{\lambda}_k) = - \ln \frac{Z(\vec{\lambda}_{k+1})}{Z(\vec{\lambda}_k)} \, .
\end{equation}
While this strategy usually results in sampling thermodynamic states whose Boltzmann distributions is very similar, thus collecting information that is to some degree redundant, some estimators such as the Multistate Bennett acceptance ratio (MBAR)~\cite{shirts2008statisticallya} can exploit similarities between states to improve the precision of the estimates by using the conformations sampled in all alchemical states $\{\vec{\lambda}_k \}$ to compute the free energy difference $\Delta f(\vec{\lambda}_i, \vec{\lambda}_{j})$ between any pair of states $i,j$  (see Section~\ref{subsec:estimators}).

@@

To ensure modeling matches experiment, we also need to accurately model and  simulate the same system -- which requires understanding what protein construct is used in the bioassay. For instance, does the X-ray structure that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer etc)~\cite{perez-benito2018predicting}? Also, were certain co-factors or partner proteins required in the bioassay? 
%
\subsection{Is my binding mode accurate?}
As also mentioned, good performance of alchemical FE calculations requires an accurate representation of the ligand binding mode, usually from a high quality X-ray crystal structure. If more than one structure is available, the modeler should pay attention to choose the most suitable. The quality of the structure can be a concern, and the reader is referred to work of Warren et al. for a detailed discussion of choosing optimal structures for structure-based modeling~\cite{warren2012essential}.

It is also useful to study the structure activity relationship and understand the expected impact in the binding site, whether side chain movement in the protein will be required, and whether there is evidence of this in any alternative X-ray structures of the same protein. Often, only one protein and water configuration is used for a series of alchemical FE calculations, so this needs to be capable of accommodating the smallest through to largest ligands in a way that allows stable and well behaved simulations. This can provide a practical limit on the alchemical changes that are feasible, though a simple work-around can be to separate compounds into sub-series for different calculations.

@@

Finally, the role of water in ligand binding is not always well understood and it can be crucial to capture the changes in binding site solvation during ligand binding. Can crystallographic waters be retained? Do they clash with some of the larger ligands used in the FEP? If so, how will the solvent be incorporated for smaller ligands that require the water – consider pre-solvation methods such as GCMC steps~\cite{michel2010prediction}. Other tools such as WaterMap or open source equivalents (SSTmap, others) can be used to define water structure for systems with no experimental evidence of water sites. Before launching large numbers of alchemical free energy calculations it is recommended to test the system using classical MD simulations and limited numbers of alchemical perturbations. Metrics such as ligand and protein RMSD and RMSF can be inspected, along with visual inspection of simulations, to ensure the system is stable and likely to be suitable for FE calculations.

@@

\paragraph{How to deal with different binding modes?}
Generally, binding modes within congeneric series are conserved~\cite{wacker2010conserved}, however, exceptions exist~\cite{brandt2011congeneric,nazare2005probing}, as discussed in more detail in Section~\ref{sec:multiple_binding_modes}. Certain  functional groups may be particularly prone to this due to symmetries or near symmetries. One such issue involves a 180 degree flip in the dihedral angle of a 6-membered aromatic ring leading to a different spatial position of ortho- or meta- substituents that otherwise should overlap within a series. The 180 degree flip of the ring may not occur enough during simulations to overcome bias due to the starting configuration. Another scenario may be equatorial and axial substituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding~\cite{kaus2015how}.
%
\paragraph{Have you considered sterioisomers and enationmers?}
Congeneric series can contain stereoisomers or enantiomers which can bind very differently, resulting in large errors if treated incorrectly. For racemates, the relative abundance of each stereoisomer is normally not known. Therefore, the experimental activity associated with just one stereoisomer/enantiomer is more uncertain. However, the modeling typically uses just the bioactive conformation that best fits the active site. Clearly this introduces potential for larger errors compared to experiment. Nevertheless, if all compounds in the congeneric series are racemic, originating from similar synthetic procedures with an expected similar abundance of stereoisomers, then the differences may cancel and the trend in calculated and observed binding energies may be robust. Despite this, we can see that care and further testing is needed in this scenario, and the quality of the predictions may suffer. Additionally, unexpected changes in what stereoisomer binds experimentally, if they occur, could pose significant challenges for modelling efforts.
%
\paragraph{Conserved binding site waters can play an important role in binding free energies}
Binding site water molecules may form water mediated protein-ligand interactions which can pose challenges whenever exchange with bulk water is slow compared to simulation timescales. This happens typically in buried binding sites. Overlaying multiple protein X-ray structures can identify conserved or additional water molecules that can be useful to include in calculations. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be useful. For examples, see HSP90 which formed part of the D3R grand challenge 2015~\cite{mey2016blinded}, A2A~\cite{brucemacdonald2018ligand}, MUP~\cite{ross2015water}, and others~\cite{michel2009energetics}.
%
\paragraph{How to align congeneric series}
Input coordinates for a congeneric series may be generated by docking calculations, or by ligand alignment using maximum common substructure algorithms. The latter tends to produce alignments that are more conserved and more consistent free energy changes across a dataset, but will struggle to yield reasonable results for relative binding free energy calculations that involve a significant binding mode rearrangement. This may also lead to steric clashes with the receptor coordinates of the reference ligand if structural rearrangements are needed to accommodate different members of the congeneric series. Small steric clashes may be resolved during subsequent simulation equilibration prior to data collection, but there is a risk that the complex relaxes to an alternative metastable state.

@@

In this case the common substructure is the benzene ring, though may be larger depending on how it is defined, as we discuss below.

@@

Rigorously, this process comprises a maximal common substructure (MCSS) search of the molecules involved to identify the common substructure -- though the parameters of the MCSS search will differ depending on whether single or dual topology calculations are planned.

@@

Some approaches have attempted to address this~\cite{clark2019relative} but a general solution is not yet in mainstream use. (at end of first sentence of this pgraph I would also cite this work from Liu, Wang and Mobley, https://pubs.acs.org/doi/abs/10.1021/acs.jcim.5b00057)

@@

The use of restraints can therefore dramatically improve sampling as interactions are weakened and removed.

@@

Alternatively, a flat-bottom restraint can be activated also in the decoupled state as long as the second restraint (e.g., a harmonic or Boresch restraint) prevents the ligand to hit the wall of the flat-bottom potential~\cite{rizzi2019sampl6}.

@@

Specifically, finite-size effects can introduce profound artifacts into computed binding free energies~\cite{}MISSING citation, in part because typical schemes for long-range electrostatics (including PME and reaction field) do not handle free energy contributions from such changes effectively or as they would be handled in a hypothetical macroscopic bulk solution~\cite{}.

@@

Many enhanced sampling techniques have been proposed to try and overcome barriers in free energy calculations and mitigate convergence to erroneous estimates~\cite{}MISSING citation. 
